An Investigation of Confusing Code Patterns in JavaScript

### 2022-12-14 background.tex
review os s04

The concept of program comprehension is central to software maintenance and feature development \cite{DBLP:conf/iwpc/TilleySP96, DBLP:journals/ibmsj/OHareT94}. The study of program comprehension in the presence of atoms of confusion was introduced by Gopstein et al. \cite{DBLP:conf/msr/GopsteinZFC18}, who defined them as small code patterns that can verifiably lead to misunderstandings. In addition, for a code pattern to be considered an atom, there must be some alternative pattern or language construct that is functionally equivalent and less likely to cause confusion. One example that occurs in many programming languages is the Change of Literal Encoding. For instance, the \clang code statement \lstinline{printf("\%d", 013);} 
often leads programmers to predict the output to be \texttt{13}, even though the correct answer is \texttt{11}. This occurs because a leading \texttt{0} in a numerical literal indicates that the number is in base \texttt{8}, a fact that is not only unknown to less experienced programmers, but misleading even for seasoned developers. After formulating a list of 19 atom candidates, Gopstein et al. \cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} were able to identify 15 code idioms and language constructs that present a statistically significant difference in comparative answer correctness when each of the 15 atoms was removed. The least confusing atom produced a 14\% boost in prediction accuracy when it was removed, whereas the most confusing one showed a 60\% accuracy increase. In a different work, Gopstein et al. \cite{DBLP:conf/msr/GopsteinZFC18} presented the results of a comprehensive study on the incidence of atoms of confusion in the wild, considering open-source projects written in \clang and \cpplang~programming languages.

>>>>

The concept of program comprehension is central to software maintenance and feature development \cite{DBLP:conf/iwpc/TilleySP96, DBLP:journals/ibmsj/OHareT94}. The study of program comprehension in the presence of atoms of confusion was introduced by Gopstein et al. \cite{DBLP:conf/msr/GopsteinZFC18}, who defined them as small code patterns that can verifiably lead to misunderstandings. In addition, for a code pattern to be considered an atom, there must be some alternative pattern or language construct that is functionally equivalent and less likely to cause confusion. One example that occurs in many programming languages is the Change of Literal Encoding. For instance, the \clang code statement \lstinline{printf("%d", 013);} 
often leads programmers to predict the output to be \texttt{13}, even though the correct answer is \texttt{11}. This occurs because a leading \texttt{0} in a numerical literal indicates that the number is in base \texttt{8}, a fact that is not only unknown to less experienced programmers, but misleading even for seasoned developers. After formulating a list of 19 atom candidates, Gopstein et al. \cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} were able to identify 15 code idioms and language constructs that present a statistically significant difference in comparative answer correctness when each of the 15 atoms was removed. The least confusing atom produced a 14\% boost in prediction accuracy when it was removed, whereas the most confusing one showed a 60\% accuracy increase. In a different work, Gopstein et al. \cite{DBLP:conf/msr/GopsteinZFC18} presented the results of a comprehensive study on the incidence of atoms of confusion in the wild, considering open-source projects written in \clang and \cpplang~programming languages.

### 2022-12-14 main.tex
Making some modifications to address the reviwers comments, in particular pertaining to the reasons for having multiple studies.
minor changes
changing the paper organization
review os s04

\newcommand{\cpplang}{C\nolinebreak\hspace{-.05em}\raisebox{.4ex}{\small\bf +}\nolinebreak\hspace{-.10em}\raisebox{.4ex}{\small\bf +}\xspace}

@@

\newcommand{\rqa}{What is the impact of the analyzed code patterns% atoms of confusion 
on the comprehension of JavaScript code?}

@@

@@

\lstdefinelanguage{JavaScript}[]{Java}{
  morekeywords={let, console.log},
  moredelim=[is][\textcolor{darkgray}]{\%\%}{\%\%},
  moredelim=[il][\textcolor{darkgray}]{§§}
}

\newcommand{\na}{10\xspace}

@@

\input{0_Abstract}

@@

\input{1_Introduction_alt}
\input{2_Background}
%\input{7_Atoms}
\input{3_Method}
\input{4_Result}
\input{discussion}
\input{5_Threats}
\input{6_Conclusion}

>>>>

\newcommand{\cpplang}{C\nolinebreak\hspace{-.05em}\raisebox{.6ex}{\tiny\bf +}\nolinebreak\hspace{-.10em}\raisebox{.6ex}{\tiny\bf +}\xspace}

@@

\newcommand{\rqa}{What is the impact of the analyzed code patterns % atoms of confusion 
 on the comprehension of JavaScript code?}

@@

\newcommand{\revised}[1]{\textcolor{blue}{#1}}

@@

\lstdefinelanguage{JavaScript}[]{Java}{
  morekeywords={var, let, console.log},
  moredelim=[is][\textcolor{darkgray}]{\%\%}{\%\%},
  moredelim=[il][\textcolor{darkgray}]{§§}
}

\lstdefinelanguage{CodeQL}[]{Java}{
  morekeywords={select, from, where, not, and}
}
\newcommand{\na}{10\xspace}

@@

\input{abstract}

@@

\input{introduction}
\input{background}
\input{methodology}
\input{s01}
\input{s02}
\input{s03}
\input{s04}
\input{discussion}
\input{threats}
\input{conclusion}

### 2022-12-14 s04.tex
presenting an overview of the studies
review os s04

\section{Prevalence Study}\label{sec:method:msr}

To understand how often the analyzed atom candidates appear in open source projects, and thus answer our third research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mine a set of GitHub open source repositories and compute the prevalence of atom candidates.

@@

We first collect the most popular GitHub repositories that are primarily written in JavaScript. We measure popularity using the project's stargazers. This metric, available through the GitHub API, represents the number of stars a project received from users of the platform. The same metric has been used in a number of previous studies as a proxy to estimate a project's popularity~\cite{gyimesi2019bugsjs,canedo:esem2020}. We then select the top 100 most popular repositories and remove projects that do not reach the first quartile of the distribution of lines of code.

@@

We mine the occurrence of atom candidates from the repositories in our curated dataset using source code queries that we wrote using the CodeQL language~\cite{moor:gttse2007}. CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as \cpplang, Java, and JavaScript). We also automate the process of running the queries and exporting the results to a format that simplifies our analysis (and also the reproduction of this study). Finally, we compute some descriptive statistics to measure the prevalence of atoms of confusion in practice.

\begin{table}[ht]
  \centering
   \caption{Some descriptive statistics about the projects used in the MSR study}
  \begin{scriptsize}
 \begin{tabular}{rrrrr}
   \toprule
                       & Min.             & Median        & Mean             & Max. \\ \midrule
 Lines of Code         & \num{5543}       & \num{36161.5} & \num{111432.33}  & \num{1278405} \\
 Num. of Forks         & \num{1244}       & \num{6078}    & \num{8906.24}    & \num{68849} \\
 Num. of Contributors  & \num{6}          & \num{285.5}   & \num{533.44}     & \num{4047} \\
 Num. of Stars         & \num{23672}      & \num{34990}   & \num{46919.51}   & \num{310935} \\

    \bottomrule
 \end{tabular}
 \end{scriptsize}
 \label{tab:projects-statistics} 
\end{table}

@@

In this section we present the results of the  software repository
mining effort, whose goal is to reveal how prevalent atom candidates are in open source JavaScript projects. 
We mined \minedprojects open source JavaScript repositories to understand how
often atoms arise in real software. As in the interview study, this study focused on atoms that were investigated in both experiments.

Similarly to previous studies~\cite{DBLP:conf/msr/GopsteinZFC18,DBLP:journals/ese/MedeirosLAAKRG19} which investigate
the prevalence of atoms of confusion in open source \clang and \cpplang projects, we found that atoms of confusion
frequently arise in JavaScript open source systems. For instance, the six most frequently found atoms
occur in at least 80\% of the projects. Considering the extremes, atom candidates Conditional Operator and Implicit Predicate were found in all repositories we mined,
while Comma Operator occurred in only 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).

@@

% latex table generated in R 4.0.3 by xtable 1.8-4 package
% Tue Apr 13 09:34:35 2021
\begin{table}[ht]
  \centering
  {\scriptsize 
\caption{Summary of atom candidate occurrences in our dataset.}
\label{tab:occurrences-summary}
\setlength\tabcolsep{2pt} % default value: 6pt
\begin{tabular}{lrr}%c}
  \toprule
Atom & Projects & Occurr./KLOC \\%& Frequency \\ 
  \midrule
Implicit Predicate & 100\% & 19.89 \\%&  intensively used \\ 
  Conditional Operator & 100\% & 10.16 \\%&  intensively used \\ 
  Omitted Curly Braces & 91.67\% & 6.61 \\%&  intensively used \\ 
  Post Increment & 90.28\% & 5.62 \\%&  intensively used \\ 
  Pre Increment & 83.33\% & 1.04 \\%& commonly used \\ 
  Assignment as Value & 81.94\% & 1.02 \\%& commonly used \\ 
  Logic as Control Flow & 56.94\% & 0.95 \\%& commonly used \\ 
  Automatic Semicolon Insertion & 33.33\% & 0.13 \\%& commonly used \\ 
  Arithmetic as Logic & 23.61\% & 0.02 \\%& rarely used \\ 
  Comma Operator & 15.28\% & 0.03 \\%& rarely used \\
   \bottomrule
\end{tabular}
  }
\end{table}

>>>>

\section{Prevalence Study}\label{sec:s04}

To understand how often the analyzed atom candidates appear in open source JavaScript projects, and thus answer our third research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mine a set of GitHub open source repositories and compute the prevalence of atom candidates.

@@

We first collect the most popular GitHub repositories that are primarily written in JavaScript. We measure popularity using the project's stargazers. This metric, available through the GitHub API, represents the number of stars a project received from users of the platform. The same metric has been used in a number of previous studies as a proxy to estimate projects' popularity~\cite{gyimesi2019bugsjs,canedo:esem2020}. We then select the top 100 most popular repositories and remove projects that do not reach the first quartile of the distribution of lines of code (excluding from our dataset small projects).

@@

We mine the occurrence of atom candidates from the repositories in our curated dataset using source code queries that we specified using the Semmle's CodeQL language~\cite{moor:gttse2007}. CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as \cpplang, Java, and JavaScript). All CodeQL queries deal with a specific atom candidate, and collect the exact location in the source code where the atoms appear. Figure~\ref{lst:codeql} shows an example of a CodeQL query that mine occurrences of the conditional operator.

Although most of our queries are simple, we have to circumvent a few corner cases to avoid false positives. For instance, we do not consider atom candidates the pre-increment occurrences that appear within
\texttt{for statement} definitions.  We automate the process of running the queries and exporting the results to a format that simplifies our analysis (and also the reproduction of this study). Finally, we compute some descriptive statistics to measure the prevalence of atoms of confusion in practice.

\begin{table}[ht]
  \centering
   \caption{Some descriptive statistics about the projects used in the MSR study}
%  \begin{scriptsize}
 \begin{tabular}{rrrrr}
   \toprule
                       & Min.             & Median        & Mean             & Max. \\ \midrule
 Lines of Code         & \num{5543}       & \num{36161.5} & \num{111432.33}  & \num{1278405} \\
 Num. of Forks         & \num{1244}       & \num{6078}    & \num{8906.24}    & \num{68849} \\
 Num. of Contributors  & \num{6}          & \num{285.5}   & \num{533.44}     & \num{4047} \\
 Num. of Stars         & \num{23672}      & \num{34990}   & \num{46919.51}   & \num{310935} \\

    \bottomrule
 \end{tabular}
 %\end{scriptsize}
 \label{tab:projects-statistics} 
\end{table}

\begin{figure}
\begin{lstlisting}[language=CodeQL]
import javascript

from ConditionalExpr e
where not e.getTopLevel().isMinified() 
select e, e.getLocation().getStartLine() as Location,e.getLocation().getFile()
\end{lstlisting}
\caption{CodeQL query that searches for Conditional Expressions.}
\label{lst:codeql}
\end{figure}

%% \begin{lstlisting}[language=CodeQL]
%% import javascript

%% from ControlStmt c
%% where not c.getTopLevel().isMinified() and
%%  (c instanceof IfStmt and not c.(IfStmt).getCondition().mayHaveBooleanValue(true)) or
%%  (c instanceof LoopStmt and  not c.(LoopStmt).getTest().mayHaveBooleanValue(true) ) or 
%%  (c instanceof WithStmt and not  c.(WithStmt).getExpr().mayHaveBooleanValue(true))  
%% select c, c.getLocation().getStartLine() as Location,c.getLocation().getFile()
%% \end{lstlisting}

@@

In this section we present the results of the  software repository
mining effort, whose goal is to reveal how prevalent atom candidates are in open source JavaScript projects. 
We mined \minedprojects open source JavaScript repositories to understand how
often atoms arise in real software. As in the interview study, this study focused on atoms that were investigated in both
experiments (detailed in Section~\ref{sec:s01} and Section~\ref{sec:s02}).

Similarly to previous studies~\cite{DBLP:conf/msr/GopsteinZFC18,DBLP:journals/ese/MedeirosLAAKRG19} which investigate
the prevalence of atoms of confusion in open source \clang and \cpplang projects, we found that atoms of confusion
frequently arise in JavaScript open source systems. For instance, the six most frequently found atoms
occur in at least 80\% of the projects. Considering the extremes, atom candidates Conditional Operator and
Implicit Predicate were found in all repositories we mined (see examples in Figure~\ref{lst:conditional-operator-sample}
and Figure~\ref{lst:implicit-predicate-sample}), while Comma Operator occurred in only 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).
Figure~\ref{lst:comma-operator-sample} shows an example of the use of the Comma Operator in the \texttt{styled-components}
project.

\begin{figure}[htb]
  \begin{lstlisting}[language=JavaScript]
function prependModifierMarker (symbol, name, dynamic) {
  return dynamic
  ? ("_p(" + name + ",\"" + symbol + "\")")
  : symbol + name // mark the event as captured
}

function concat (a, b) {
  return a ? b ? (a + ' ' + b) : a : (b || '')
}
  \end{lstlisting}
  \caption{Examples of the Conditional Operator atom present in the \texttt{vue} project (urls: shorturl.at/bJPS4 and shorturl.at/tJOS8)}
   \label{lst:conditional-operator-sample}
\end{figure}

\begin{figure}[htb]
  \begin{lstlisting}[language=JavaScript]
React.useEffect(() => {
  if (props.config.selectedLines) { // Implicit Predicate
    dispatch({
      type: 'MULTILINE',
      selectedLines: props.config.selectedLines,
    })
  }
}, [props.config.selectedLines])
  \end{lstlisting}
  \caption{Example of the Implicit Predicate atom present in the \texttt{carbon} project (url: shorturl.at/aloqU)}
  \label{lst:implicit-predicate-sample}
\end{figure}

\begin{figure}[htb]
  \begin{lstlisting}[language=JavaScript]
var props = ((_props = {}),
 (_props[SC_ATTR] = stringifyNames(names)),
 (_props[SC_VERSION_ATTR] = '4.3.1'),
 _props);
  \end{lstlisting}
  \caption{Example of the Comma Operator atom present in the \texttt{styled-components} project (url: shorturl.at/sCKM0)}
  \label{lst:comma-operator-sample}
\end{figure}

@@

% latex table generated in R 4.0.3 by xtable 1.8-4 package
% Tue Apr 13 09:34:35 2021
\begin{table}[ht]
  \centering
\caption{Summary of atom candidate occurrences in our dataset.}
\label{tab:occurrences-summary}
\setlength\tabcolsep{2pt} % default value: 6pt
\begin{tabular}{lrr}%c}
  \toprule
Atom & Projects & Occurr./KLOC \\%& Frequency \\ 
  \midrule
Implicit Predicate & 100\% & 19.89 \\%&  intensively used \\ 
  Conditional Operator & 100\% & 10.16 \\%&  intensively used \\ 
  Omitted Curly Braces & 91.67\% & 6.61 \\%&  intensively used \\ 
  Post Increment & 90.28\% & 5.62 \\%&  intensively used \\ 
  Pre Increment & 83.33\% & 1.04 \\%& commonly used \\ 
  Assignment as Value & 81.94\% & 1.02 \\%& commonly used \\ 
  Logic as Control Flow & 56.94\% & 0.95 \\%& commonly used \\ 
  Automatic Semicolon Insertion & 33.33\% & 0.13 \\%& commonly used \\ 
  Arithmetic as Logic & 23.61\% & 0.02 \\%& rarely used \\ 
  Comma Operator & 15.28\% & 0.03 \\%& rarely used \\
   \bottomrule
\end{tabular}

\end{table}
